---
title: "BSA Data Journalism: Who is The Most 'Like Mike'"
output: html_document
date: "2023-02-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Player 1: Michael Jordan
- prime years: 91 - 93, 96 - 98
- often referenced as the best 6 years of basketball ever
- arguably the greatest prime in NBA history
- two 3 peats in the NBA finals
```{r}
mj_basic <- read_csv('mj.csv')
str(mj_basic) # seeing the data types of the variables
dim(mj_basic) # checking the dimensions in order to subset columns in vars()
mj_basic_prime <- mj_basic %>% filter(Age %in% c(27:29, 32:34)) %>% mutate_at(vars(6:30), as.numeric)
mj_basic_prime
```

Player 2: Dwyane Wade
An explosive shooting guard and mid range threat.
- prime years: 2006 - 2007, 2009 - 2011, 2013 
- 2005 - 2006 saw his rise as a young star with his first title in 2006 as the main option for the Heat
- 2009 - 2011 saw him enter his peak as he won the scoring title in 2009
- Due to injuries in the 2012 season, his last year for consideration will be 2013 where he won another championship while greatly contributing to his team's success.
```{r}
dwade_basic <- read_csv('dwadebasic.csv')
dwade_basic_prime <- dwade_basic %>% filter(Age %in% c(24:25, 27:29, 31))
dwade_basic_prime
```

Player 3: DeMar DeRozan
Highly athletic shooting guard, Jordan-like shot selection, mid range specialist, explosive finishing.
- prime years: 2016 - 2019, 2022 - 2023
- 14 - 17 saw his rise to stardom as a key component to the success of the Raptors
- After having slower seasons with the Spurs, saw new explosive success as the primary threat for the Bulls starting from last season. 
```{r}
derozan_basic <- read_csv('derozanbasic.csv')
derozan_basic_prime <- derozan_basic %>% filter(Age %in% c(26:29, 32:33))
derozan_basic_prime
```

Player 4: Kobe Bryant
Similar drive to win and be the best, Jordan-like shot selection, threat from anywhere on the court, highly athletic.
```{r}
kobe_basic <- read_csv('kobebasic.csv')
kobe_basic_prime <- kobe_basic %>% filter (Age %in% c(22:23, 27:28, 30:31))
kobe_basic_prime
```


Player 5: Lebron James
Considered one of, if not, the biggest threat to being considered the GOAT. Seen as a player of comparable dominance over the league.
- prime years: 2009 - 2014
- often seen as his most dominant hold on the league
- won 4 MVPS, 2 Championships, 2 FMVPS
- can argue any 6 year stretch of his career can be his prime years
```{r}
lebron_basic <- read_csv('lebronbasic.csv')
lebron_basic_prime <- lebron_basic %>% filter (Age %in% c(24:29))
lebron_basic_prime
```


Player 6: Kawhi Leondard
Defensive pressure, mid range specialist, Jordan-like shot selection.
- prime years: 2016 - 2021
- saw his rise to stardom on the spurs
- all of his 5 ASG appearances in this interval
- where he won a ring with Toronto and won the FMVP
- won FMVP with SAS in 2014
```{r}
kawhi_basic <- read_csv('kawhibasic.csv')
str(kawhi_basic) # seeing the data types of the variables
dim(kawhi_basic) # checking the dimensions in order to subset columns in vars()
kawhi_basic_prime <- kawhi_basic %>% filter(Age %in% c(24:29)) %>% mutate_at(vars(6:30), as.numeric)
kawhi_basic_prime
```

