---
title: "BSA Data Journalism: Who is The Most 'Like Mike'"
output: html_document
date: "2023-02-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Player 1: Michael Jordan
- prime years: 91 - 93, 96 - 98
- often referenced as the best 6 years of basketball ever
- arguably the greatest prime in NBA history
- two 3 peats in the NBA finals
```{r}
mj_basic <- read_csv('mj.csv')
str(mj_basic) # seeing the data types of the variables
dim(mj_basic) # checking the dimensions in order to subset columns in vars()
mj_basic_prime <- mj_basic %>% filter(Age %in% c(27:29, 32:34)) %>% mutate_at(vars(6:30), as.numeric)
mj_basic_prime$PrimeYear <- 1:6
mj_basic_prime
```

Player 2: Dwyane Wade
An explosive shooting guard and mid range threat.
- prime years: 2006 - 2007, 2009 - 2011, 2013 
- 2005 - 2006 saw his rise as a young star with his first title in 2006 as the main option for the Heat
- 2009 - 2011 saw him enter his peak as he won the scoring title in 2009
- Due to injuries in the 2012 season, his last year for consideration will be 2013 where he won another championship while greatly contributing to his team's success.
```{r}
dwade_basic <- read_csv('dwadebasic.csv')
dwade_basic_prime <- dwade_basic %>% filter(Age %in% c(24:25, 27:29, 31))
dwade_basic_prime$PrimeYear <- 1:6
dwade_basic_prime
```

Player 3: DeMar DeRozan
Highly athletic shooting guard, Jordan-like shot selection, mid range specialist, explosive finishing.
- prime years: 2016 - 2019, 2022 - 2023
- 14 - 17 saw his rise to stardom as a key component to the success of the Raptors
- After having slower seasons with the Spurs, saw new explosive success as the primary threat for the Bulls starting from last season. 
```{r}
derozan_basic <- read_csv('derozanbasic.csv')
derozan_basic_prime <- derozan_basic %>% filter(Age %in% c(26:29, 32:33))
derozan_basic_prime$PrimeYear <- 1:6
derozan_basic_prime
```

Player 4: Kobe Bryant
Similar drive to win and be the best, Jordan-like shot selection, threat from anywhere on the court, highly athletic.
```{r}
kobe_basic <- read_csv('kobebasic.csv')
kobe_basic_prime <- kobe_basic %>% filter (Age %in% c(22:23, 27:28, 30:31))
kobe_basic_prime$PrimeYear <- 1:6
kobe_basic_prime
```


Player 5: LeBron James
Considered one of, if not, the biggest threat to being considered the GOAT. Seen as a player of comparable dominance over the league.
- prime years: 2009 - 2014
- often seen as his most dominant hold on the league
- won 4 MVPS, 2 Championships, 2 FMVPS
- can argue any 6 year stretch of his career can be his prime years
```{r}
lebron_basic <- read_csv('lebronbasic.csv')
lebron_basic_prime <- lebron_basic %>% filter (Age %in% c(24:29))
lebron_basic_prime$PrimeYear <- 1:6
lebron_basic_prime
```


Player 6: Kawhi Leonard
Defensive pressure, mid range specialist, Jordan-like shot selection.
- prime years: 2016 - 2021
- saw his rise to stardom on the spurs
- all of his 5 ASG appearances in this interval
- where he won a ring with Toronto and won the FMVP
- won FMVP with SAS in 2014
```{r}
kawhi_basic <- read_csv('kawhibasic.csv')
str(kawhi_basic) # seeing the data types of the variables
dim(kawhi_basic) # checking the dimensions in order to subset columns in vars()
kawhi_basic_prime <- kawhi_basic %>% filter(Age %in% c(24:29)) %>% mutate_at(vars(6:30), as.numeric)
kawhi_basic_prime$PrimeYear <- 1:6
kawhi_basic_prime
```


Methodology: Similarity Stat
- in order to compare these players we must implement a 'similarity stat'
- this will take into account both basic and advances stats
  - PPG, RPG, APG, FG%, SPG, Box +/-, Winshares, Shot selection, O/D Rating
  - a percentage of similarity for each of these statistics will be taken, then standardized to a number between 1 and 10
  - all 10 categories will then be added up and a player will recieve a score between 1 and 100
  
1. Dwayne Wade
  
```{r}
ggplot() +
  geom_line(data = mj_basic_prime, aes(x = PrimeYear, y = PTS), color = "red") + 
  geom_line(data = dwade_basic_prime, aes(x = PrimeYear, y = PTS)) + 
  geom_line(data = derozan_basic_prime, aes(x = PrimeYear, y = PTS), color = "gray") + 
  geom_line(data = kobe_basic_prime, aes(x = PrimeYear, y = PTS), color = "purple") + 
  geom_line(data = lebron_basic_prime, aes(x = PrimeYear, y = PTS), color = "yellow") +
  geom_line(data = kawhi_basic_prime, aes(x = PrimeYear, y = PTS), color = "blue") +
  geom_point(data = mj_basic_prime, aes(x = PrimeYear, y = PTS), color = "red") + 
  geom_point(data = dwade_basic_prime, aes(x = PrimeYear, y = PTS)) +
  geom_point(data = derozan_basic_prime, aes(x = PrimeYear, y = PTS), color = "gray") + 
  geom_point(data = kobe_basic_prime, aes(x = PrimeYear, y = PTS), color = "purple") + 
  geom_point(data = lebron_basic_prime, aes(x = PrimeYear, y = PTS), color = "yellow") +
  geom_point(data = kawhi_basic_prime, aes(x = PrimeYear, y = PTS), color = "blue")
```

- from this visualization we see Dwade has a similar trend
- Lebron has similar consistency
- Kobe peaked the highest
  
```{r}
appg <- tribble(
~Player, ~Avg_PPG,
#Sad?/12 months/an account
"Michael Jordan", mean(mj_basic_prime$PTS),
"Dwyane Wade", mean(dwade_basic_prime$PTS),
"DeMar DeRozan", mean(derozan_basic_prime$PTS),
"Kobe Bryant", mean(kobe_basic_prime$PTS),
"LeBron James", mean(lebron_basic_prime$PTS),
"Kawhi Leonard", mean(kawhi_basic_prime$PTS),
)

appg %>% ggplot() +
  geom_bar(aes(x = Player, y = Avg_PPG, color = Player, fill = Player), stat = 'identity')
```

Criteria:

Point Differential        Score

0 - 1                     10

1.1 - 2                    9

...                         ...



```{r}
appg$Avg_PPG - mean(mj_basic_prime$PTS)
```

SCORES:
Wade: 6
Derozan: 5
Kobe: 9
Lebron: 8
Kawhi: 4


```{r}
ggplot() +
  geom_line(data = mj_basic_prime, aes(x = PrimeYear, y = STL), color = "red") + 
  geom_line(data = dwade_basic_prime, aes(x = PrimeYear, y = STL)) + 
  geom_line(data = derozan_basic_prime, aes(x = PrimeYear, y = STL), color = "gray") + 
  geom_line(data = kobe_basic_prime, aes(x = PrimeYear, y = STL), color = "purple") + 
  geom_line(data = lebron_basic_prime, aes(x = PrimeYear, y = STL), color = "yellow") +
  geom_line(data = kawhi_basic_prime, aes(x = PrimeYear, y = STL), color = "blue") +
  geom_point(data = mj_basic_prime, aes(x = PrimeYear, y = STL), color = "red") + 
  geom_point(data = dwade_basic_prime, aes(x = PrimeYear, y = STL)) +
  geom_point(data = derozan_basic_prime, aes(x = PrimeYear, y = STL), color = "gray") + 
  geom_point(data = kobe_basic_prime, aes(x = PrimeYear, y = STL), color = "purple") + 
  geom_point(data = lebron_basic_prime, aes(x = PrimeYear, y = STL), color = "yellow") +
  geom_point(data = kawhi_basic_prime, aes(x = PrimeYear, y = STL), color = "blue")
```

```{r}
aspg <- tribble(
~Player, ~Avg_SPG,
#Sad?/12 months/an account
"Michael Jordan", mean(mj_basic_prime$STL),
"Dwyane Wade", mean(dwade_basic_prime$STL),
"DeMar DeRozan", mean(derozan_basic_prime$STL),
"Kobe Bryant", mean(kobe_basic_prime$STL),
"LeBron James", mean(lebron_basic_prime$STL),
"Kawhi Leonard", mean(kawhi_basic_prime$STL),
)

aspg %>% ggplot() +
  geom_bar(aes(x = Player, y = Avg_SPG, color = Player, fill = Player), stat = 'identity')
```


```{r}
aspg$Avg_SPG - mean(mj_basic_prime$STL)
```


```{r}
ggplot() +
  geom_line(data = mj_basic_prime, aes(x = PrimeYear, y = `FG%`), color = "red") + 
  geom_line(data = dwade_basic_prime, aes(x = PrimeYear, y = `FG%`)) + 
  geom_line(data = derozan_basic_prime, aes(x = PrimeYear, y = `FG%`), color = "gray") + 
  geom_line(data = kobe_basic_prime, aes(x = PrimeYear, y = `FG%`), color = "purple") + 
  geom_line(data = lebron_basic_prime, aes(x = PrimeYear, y = `FG%`), color = "yellow") +
  geom_line(data = kawhi_basic_prime, aes(x = PrimeYear, y = `FG%`), color = "blue") +
  geom_point(data = mj_basic_prime, aes(x = PrimeYear, y = `FG%`), color = "red") + 
  geom_point(data = dwade_basic_prime, aes(x = PrimeYear, y = `FG%`)) +
  geom_point(data = derozan_basic_prime, aes(x = PrimeYear, y = `FG%`), color = "gray") + 
  geom_point(data = kobe_basic_prime, aes(x = PrimeYear, y = `FG%`), color = "purple") + 
  geom_point(data = lebron_basic_prime, aes(x = PrimeYear, y = `FG%`), color = "yellow") +
  geom_point(data = kawhi_basic_prime, aes(x = PrimeYear, y = `FG%`), color = "blue")
```


```{r}
afgp <- tribble(
~Player, ~`Avg_FG%`,
#Sad?/12 months/an account
"Michael Jordan", mean(mj_basic_prime$`FG%`),
"Dwyane Wade", mean(dwade_basic_prime$`FG%`),
"DeMar DeRozan", mean(derozan_basic_prime$`FG%`),
"Kobe Bryant", mean(kobe_basic_prime$`FG%`),
"LeBron James", mean(lebron_basic_prime$`FG%`),
"Kawhi Leonard", mean(kawhi_basic_prime$`FG%`),
)

afgp %>% ggplot() +
  geom_bar(aes(x = Player, y = `Avg_FG%`, color = Player, fill = Player), stat = 'identity')
```

```{r}
afgp$`Avg_FG%` - mean(mj_basic_prime$`FG%`)
```
