---
title: "BSA Data Journalism: Who is The Most 'Like Mike'"
output: html_document
date: "2023-02-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Player 1: Michael Jordan
- prime years: 91 - 93, 96 - 98
- often referenced as the best 6 years of basketball ever
- arguably the greatest prime in NBA history
- two 3 peats in the NBA finals
```{r}
mj_basic <- read_csv('mj.csv')
str(mj_basic) # seeing the data types of the variables
dim(mj_basic) # checking the dimensions in order to subset columns in vars()
mj_basic_prime <- mj_basic %>% filter(Age %in% c(27:29, 32:34)) %>% mutate_at(vars(6:30), as.numeric)
mj_basic_prime$PrimeYear <- 1:6
mj_basic_prime
```

Player 2: Dwyane Wade
An explosive shooting guard and mid range threat.
- prime years: 2006 - 2007, 2009 - 2011, 2013 
- 2005 - 2006 saw his rise as a young star with his first title in 2006 as the main option for the Heat
- 2009 - 2011 saw him enter his peak as he won the scoring title in 2009
- Due to injuries in the 2012 season, his last year for consideration will be 2013 where he won another championship while greatly contributing to his team's success.
```{r}
dwade_basic <- read_csv('dwadebasic.csv')
dwade_basic_prime <- dwade_basic %>% filter(Age %in% c(24:25, 27:29, 31))
dwade_basic_prime$PrimeYear <- 1:6
dwade_basic_prime
```

Player 3: DeMar DeRozan
Highly athletic shooting guard, Jordan-like shot selection, mid range specialist, explosive finishing.
- prime years: 2016 - 2019, 2022 - 2023
- 14 - 17 saw his rise to stardom as a key component to the success of the Raptors
- After having slower seasons with the Spurs, saw new explosive success as the primary threat for the Bulls starting from last season. 
```{r}
derozan_basic <- read_csv('derozanbasic.csv')
derozan_basic_prime <- derozan_basic %>% filter(Age %in% c(26:29, 32:33))
derozan_basic_prime$PrimeYear <- 1:6
derozan_basic_prime
```

Player 4: Kobe Bryant
Similar drive to win and be the best, Jordan-like shot selection, threat from anywhere on the court, highly athletic.
```{r}
kobe_basic <- read_csv('kobebasic.csv')
kobe_basic_prime <- kobe_basic %>% filter (Age %in% c(22:23, 27:28, 30:31))
kobe_basic_prime$PrimeYear <- 1:6
kobe_basic_prime
```


Player 5: LeBron James
Considered one of, if not, the biggest threat to being considered the GOAT. Seen as a player of comparable dominance over the league.
- prime years: 2009 - 2014
- often seen as his most dominant hold on the league
- won 4 MVPS, 2 Championships, 2 FMVPS
- can argue any 6 year stretch of his career can be his prime years
```{r}
lebron_basic <- read_csv('lebronbasic.csv')
lebron_basic_prime <- lebron_basic %>% filter (Age %in% c(24:29))
lebron_basic_prime$PrimeYear <- 1:6
lebron_basic_prime
```


Player 6: Kawhi Leonard
Defensive pressure, mid range specialist, Jordan-like shot selection.
- prime years: 2016 - 2021
- saw his rise to stardom on the spurs
- all of his 5 ASG appearances in this interval
- where he won a ring with Toronto and won the FMVP
- won FMVP with SAS in 2014
```{r}
kawhi_basic <- read_csv('kawhibasic.csv')
str(kawhi_basic) # seeing the data types of the variables
dim(kawhi_basic) # checking the dimensions in order to subset columns in vars()
kawhi_basic_prime <- kawhi_basic %>% filter(Age %in% c(24:29)) %>% mutate_at(vars(6:30), as.numeric)
kawhi_basic
kawhi_basic_prime$PrimeYear <- 1:6
kawhi_basic_prime
```


Methodology: Similarity Stat
- in order to compare these players we must implement a 'similarity stat'
- this will take into account both basic and advances stats
  - PPG, RPG, APG, FG%, SPG, Box +/-, Winshares, Shot selection, O/D Rating
  - a percentage of similarity for each of these statistics will be taken, then standardized to a number between 1 and 10
  - all 10 categories will then be added up and a player will recieve a score between 1 and 100
  
1. Dwayne Wade
  
```{r}
ggplot() +
  geom_line(data = mj_basic_prime, aes(x = PrimeYear, y = PTS), color = "red") + 
  geom_line(data = dwade_basic_prime, aes(x = PrimeYear, y = PTS), color = "black") + 
  geom_line(data = derozan_basic_prime, aes(x = PrimeYear, y = PTS, color = "gray")) + 
  geom_line(data = kobe_basic_prime, aes(x = PrimeYear, y = PTS), color = "purple") + 
  geom_line(data = lebron_basic_prime, aes(x = PrimeYear, y = PTS), color = "yellow") +
  geom_line(data = kawhi_basic_prime, aes(x = PrimeYear, y = PTS), color = "blue") +
  geom_point(data = mj_basic_prime, aes(x = PrimeYear, y = PTS), color = "red") + 
  geom_point(data = dwade_basic_prime, aes(x = PrimeYear, y = PTS), color = "black") +
  geom_point(data = derozan_basic_prime, aes(x = PrimeYear, y = PTS, color = "gray")) + 
  geom_point(data = kobe_basic_prime, aes(x = PrimeYear, y = PTS), color = "purple") + 
  geom_point(data = lebron_basic_prime, aes(x = PrimeYear, y = PTS), color = "yellow") +
  geom_point(data = kawhi_basic_prime, aes(x = PrimeYear, y = PTS), color = "blue") +
  scale_color_manual(name='Player',
                     breaks=c('Jordan', 'Wade', 'DeRozan', 'Bryant', 'James', 'Leonard'),
                     values=c('Jordan' = "red", 'Wade' = "black", 'DeRozan' = "gray", 'Bryant'="purple", 'James'="yellow", 'Leonard'="blue")) +
  theme(legend.title=element_text(size=12), legend.text=element_text(size=10)) 
```

- from this visualization we see Dwade has a similar trend
- Lebron has similar consistency
- Kobe peaked the highest
  
```{r}
appg <- tribble(
~Player, ~Avg_PPG,
#Sad?/12 months/an account
"Michael Jordan", mean(mj_basic_prime$PTS),
"Dwyane Wade", mean(dwade_basic_prime$PTS),
"DeMar DeRozan", mean(derozan_basic_prime$PTS),
"Kobe Bryant", mean(kobe_basic_prime$PTS),
"LeBron James", mean(lebron_basic_prime$PTS),
"Kawhi Leonard", mean(kawhi_basic_prime$PTS),
)

appg %>% ggplot() +
  geom_bar(aes(x = Player, y = Avg_PPG, color = Player, fill = Player), stat = 'identity')
```


```{r}
appg$Avg_PPG - mean(mj_basic_prime$PTS)
```


```{r}
ggplot() +
  geom_line(data = mj_basic_prime, aes(x = PrimeYear, y = STL), color = "red") + 
  geom_line(data = dwade_basic_prime, aes(x = PrimeYear, y = STL), color = "black") + 
  geom_line(data = derozan_basic_prime, aes(x = PrimeYear, y = STL, color = "gray")) + 
  geom_line(data = kobe_basic_prime, aes(x = PrimeYear, y = STL), color = "purple") + 
  geom_line(data = lebron_basic_prime, aes(x = PrimeYear, y = STL), color = "yellow") +
  geom_line(data = kawhi_basic_prime, aes(x = PrimeYear, y = STL), color = "blue") +
  geom_point(data = mj_basic_prime, aes(x = PrimeYear, y = STL), color = "red") + 
  geom_point(data = dwade_basic_prime, aes(x = PrimeYear, y = STL), color = "black") +
  geom_point(data = derozan_basic_prime, aes(x = PrimeYear, y = STL, color = "gray")) + 
  geom_point(data = kobe_basic_prime, aes(x = PrimeYear, y = STL), color = "purple") + 
  geom_point(data = lebron_basic_prime, aes(x = PrimeYear, y = STL), color = "yellow") +
  geom_point(data = kawhi_basic_prime, aes(x = PrimeYear, y = STL), color = "blue") +
  scale_color_manual(name='Player',
                     breaks=c('Jordan', 'Wade', 'DeRozan', 'Bryant', 'James', 'Leonard'),
                     values=c('Jordan' = "red", 'Wade' = "black", 'DeRozan' = "gray", 'Bryant'="purple", 'James'="yellow", 'Leonard'="blue")) +
  theme(legend.title=element_text(size=12), legend.text=element_text(size=10)) 
```

```{r}
aspg <- tribble(
~Player, ~Avg_SPG,
#Sad?/12 months/an account
"Michael Jordan", mean(mj_basic_prime$STL),
"Dwyane Wade", mean(dwade_basic_prime$STL),
"DeMar DeRozan", mean(derozan_basic_prime$STL),
"Kobe Bryant", mean(kobe_basic_prime$STL),
"LeBron James", mean(lebron_basic_prime$STL),
"Kawhi Leonard", mean(kawhi_basic_prime$STL),
)

aspg %>% ggplot() +
  geom_bar(aes(x = Player, y = Avg_SPG, color = Player, fill = Player), stat = 'identity')
```


```{r}
aspg$Avg_SPG - mean(mj_basic_prime$STL)
```


```{r}
ggplot() +
  geom_line(data = mj_basic_prime, aes(x = PrimeYear, y = `FG%`), color = "red") + 
  geom_line(data = dwade_basic_prime, aes(x = PrimeYear, y = `FG%`), color = "black") + 
  geom_line(data = derozan_basic_prime, aes(x = PrimeYear, y = `FG%`, color = "gray")) + 
  geom_line(data = kobe_basic_prime, aes(x = PrimeYear, y = `FG%`), color = "purple") + 
  geom_line(data = lebron_basic_prime, aes(x = PrimeYear, y = `FG%`), color = "yellow") +
  geom_line(data = kawhi_basic_prime, aes(x = PrimeYear, y = `FG%`), color = "blue") +
  geom_point(data = mj_basic_prime, aes(x = PrimeYear, y = `FG%`), color = "red") + 
  geom_point(data = dwade_basic_prime, aes(x = PrimeYear, y = `FG%`), color = "black") +
  geom_point(data = derozan_basic_prime, aes(x = PrimeYear, y = `FG%`, color = "gray")) + 
  geom_point(data = kobe_basic_prime, aes(x = PrimeYear, y = `FG%`), color = "purple") + 
  geom_point(data = lebron_basic_prime, aes(x = PrimeYear, y = `FG%`), color = "yellow") +
  geom_point(data = kawhi_basic_prime, aes(x = PrimeYear, y = `FG%`), color = "blue") +
  scale_color_manual(name='Player',
                     breaks=c('Jordan', 'Wade', 'DeRozan', 'Bryant', 'James', 'Leonard'),
                     values=c('Jordan' = "red", 'Wade' = "black", 'DeRozan' = "gray", 'Bryant'="purple", 'James'="yellow", 'Leonard'="blue")) +
  theme(legend.title=element_text(size=12), legend.text=element_text(size=10)) 
```


```{r}
afgp <- tribble(
~Player, ~`Avg_FG%`,
#Sad?/12 months/an account
"Michael Jordan", mean(mj_basic_prime$`FG%`),
"Dwyane Wade", mean(dwade_basic_prime$`FG%`),
"DeMar DeRozan", mean(derozan_basic_prime$`FG%`),
"Kobe Bryant", mean(kobe_basic_prime$`FG%`),
"LeBron James", mean(lebron_basic_prime$`FG%`),
"Kawhi Leonard", mean(kawhi_basic_prime$`FG%`),
)

afgp %>% ggplot() +
  geom_bar(aes(x = Player, y = `Avg_FG%`, color = Player, fill = Player), stat = 'identity')
```

```{r}
afgp$`Avg_FG%` - mean(mj_basic_prime$`FG%`)
```



```{r}
library(httr)
# Installing Packages
install.packages("ClusterR")
install.packages("cluster")
  
# Loading package
library(ClusterR)
library(cluster)

headers = c(
  `Connection` = 'keep-alive',
  `Accept` = 'application/json, text/plain, */*',
  `x-nba-stats-token` = 'true',
  `X-NewRelic-ID` = 'VQECWF5UChAHUlNTBwgBVw==',
  `User-Agent` = 'Mozilla/5.0 (Macintosh; Intel Mac OS X 10_14_6) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/78.0.3904.87 Safari/537.36',
  `x-nba-stats-origin` = 'stats',
  `Sec-Fetch-Site` = 'same-origin',
  `Sec-Fetch-Mode` = 'cors',
  `Referer` = 'https://stats.nba.com/players/leaguedashplayerbiostats/',
  `Accept-Encoding` = 'gzip, deflate, br',
  `Accept-Language` = 'en-US,en;q=0.9'
)

url <- "https://stats.nba.com/stats/shotchartdetail?AheadBehind=&CFID=33&CFPARAMS=2019-20&ClutchTime=&Conference=&ContextFilter=&ContextMeasure=FGA&DateFrom=&DateTo=&Division=&EndPeriod=10&EndRange=28800&GROUP_ID=&GameEventID=&GameID=&GameSegment=&GroupID=&GroupMode=&GroupQuantity=5&LastNGames=0&LeagueID=00&Location=&Month=0&OnOff=&OpponentTeamID=0&Outcome=&PORound=0&Period=0&PlayerID=893&PlayerID1=&PlayerID2=&PlayerID3=&PlayerID4=&PlayerID5=&PlayerPosition=&PointDiff=&Position=&RangeType=0&RookieYear=&Season=1996-97&SeasonSegment=&SeasonType=Regular+Season&ShotClockRange=&StartPeriod=1&StartRange=0&StarterBench=&TeamID=0&VsConference=&VsDivision=&VsPlayerID1=&VsPlayerID2=&VsPlayerID3=&VsPlayerID4=&VsPlayerID5=&VsTeamID="

res <- httr::GET(url = url, httr::add_headers(.headers=headers))
json_resp <- jsonlite::fromJSON(content(res, "text"))

df <- data.frame(json_resp$resultSets$rowSet[1])
colnames(df) <- json_resp[["resultSets"]][["headers"]][[1]]

df$LOC_X <- as.numeric(as.character(df$LOC_X))
df$LOC_Y <- as.numeric(as.character(df$LOC_Y))

df %>% 
  ggplot(aes(x = LOC_X, y = LOC_Y)) + 
  geom_point() +
  coord_equal()
jordan <- mean(as.numeric(df$SHOT_DISTANCE))
```


```{r}
ggplot(df, aes(x=LOC_X, y=LOC_Y)) +
  geom_point() +
  theme_bw() + xlim(-250, 250) + ylim(-50, 450) + ggtitle("Michael Jordan FGA: 1996 - 1997")
```




```{r}
headers = c(
  `Connection` = 'keep-alive',
  `Accept` = 'application/json, text/plain, */*',
  `x-nba-stats-token` = 'true',
  `X-NewRelic-ID` = 'VQECWF5UChAHUlNTBwgBVw==',
  `User-Agent` = 'Mozilla/5.0 (Macintosh; Intel Mac OS X 10_14_6) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/78.0.3904.87 Safari/537.36',
  `x-nba-stats-origin` = 'stats',
  `Sec-Fetch-Site` = 'same-origin',
  `Sec-Fetch-Mode` = 'cors',
  `Referer` = 'https://stats.nba.com/players/leaguedashplayerbiostats/',
  `Accept-Encoding` = 'gzip, deflate, br',
  `Accept-Language` = 'en-US,en;q=0.9'
)

url <- "https://stats.nba.com/stats/shotchartdetail?AheadBehind=&CFID=33&CFPARAMS=2019-20&ClutchTime=&Conference=&ContextFilter=&ContextMeasure=FGA&DateFrom=&DateTo=&Division=&EndPeriod=10&EndRange=28800&GROUP_ID=&GameEventID=&GameID=&GameSegment=&GroupID=&GroupMode=&GroupQuantity=5&LastNGames=0&LeagueID=00&Location=&Month=0&OnOff=&OpponentTeamID=0&Outcome=&PORound=0&Period=0&PlayerID=2548&PlayerID1=&PlayerID2=&PlayerID3=&PlayerID4=&PlayerID5=&PlayerPosition=&PointDiff=&Position=&RangeType=0&RookieYear=&Season=2008-09&SeasonSegment=&SeasonType=Regular+Season&ShotClockRange=&StartPeriod=1&StartRange=0&StarterBench=&TeamID=0&VsConference=&VsDivision=&VsPlayerID1=&VsPlayerID2=&VsPlayerID3=&VsPlayerID4=&VsPlayerID5=&VsTeamID="

res <- httr::GET(url = url, httr::add_headers(.headers=headers))
json_resp <- jsonlite::fromJSON(content(res, "text"))

df <- data.frame(json_resp$resultSets$rowSet[1])
colnames(df) <- json_resp[["resultSets"]][["headers"]][[1]]

df$LOC_X <- as.numeric(as.character(df$LOC_X))
df$LOC_Y <- as.numeric(as.character(df$LOC_Y))

df %>% 
  ggplot(aes(x = LOC_X, y = LOC_Y)) + 
  geom_point() +
  coord_equal() + xlim(-200, 200) + ylim(0, 450)

wade<- mean(as.numeric(df$SHOT_DISTANCE))
```

```{r}
ggplot(df, aes(x=LOC_X, y=LOC_Y) ) +
  geom_point() +
  theme_bw() + xlim(-250, 250) + ylim(-50, 450) + ggtitle("Dwyane Wade FGA: 2008 - 2009")
```


```{r}
headers = c(
  `Connection` = 'keep-alive',
  `Accept` = 'application/json, text/plain, */*',
  `x-nba-stats-token` = 'true',
  `X-NewRelic-ID` = 'VQECWF5UChAHUlNTBwgBVw==',
  `User-Agent` = 'Mozilla/5.0 (Macintosh; Intel Mac OS X 10_14_6) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/78.0.3904.87 Safari/537.36',
  `x-nba-stats-origin` = 'stats',
  `Sec-Fetch-Site` = 'same-origin',
  `Sec-Fetch-Mode` = 'cors',
  `Referer` = 'https://stats.nba.com/players/leaguedashplayerbiostats/',
  `Accept-Encoding` = 'gzip, deflate, br',
  `Accept-Language` = 'en-US,en;q=0.9'
)

url <- "https://stats.nba.com/stats/shotchartdetail?AheadBehind=&CFID=33&CFPARAMS=2019-20&ClutchTime=&Conference=&ContextFilter=&ContextMeasure=FGA&DateFrom=&DateTo=&Division=&EndPeriod=10&EndRange=28800&GROUP_ID=&GameEventID=&GameID=&GameSegment=&GroupID=&GroupMode=&GroupQuantity=5&LastNGames=0&LeagueID=00&Location=&Month=0&OnOff=&OpponentTeamID=0&Outcome=&PORound=0&Period=0&PlayerID=201942&PlayerID1=&PlayerID2=&PlayerID3=&PlayerID4=&PlayerID5=&PlayerPosition=&PointDiff=&Position=&RangeType=0&RookieYear=&Season=2016-17&SeasonSegment=&SeasonType=Regular+Season&ShotClockRange=&StartPeriod=1&StartRange=0&StarterBench=&TeamID=0&VsConference=&VsDivision=&VsPlayerID1=&VsPlayerID2=&VsPlayerID3=&VsPlayerID4=&VsPlayerID5=&VsTeamID="

res <- httr::GET(url = url, httr::add_headers(.headers=headers))
json_resp <- jsonlite::fromJSON(content(res, "text"))

df <- data.frame(json_resp$resultSets$rowSet[1])
colnames(df) <- json_resp[["resultSets"]][["headers"]][[1]]

df$LOC_X <- as.numeric(as.character(df$LOC_X))
df$LOC_Y <- as.numeric(as.character(df$LOC_Y))

df %>% 
  ggplot(aes(x = LOC_X, y = LOC_Y)) + 
  geom_point() +
  coord_equal() + xlim(-200, 200) + ylim(0, 450)

derozan <- mean(as.numeric(df$SHOT_DISTANCE))
```


```{r}
ggplot(df, aes(x=LOC_X, y=LOC_Y) ) +
  geom_point() +
  theme_bw() + xlim(-250, 250) + ylim(-50, 450) + ggtitle("DeMar DeRozan FGA: 2016 - 2017")
```



```{r}
headers = c(
  `Connection` = 'keep-alive',
  `Accept` = 'application/json, text/plain, */*',
  `x-nba-stats-token` = 'true',
  `X-NewRelic-ID` = 'VQECWF5UChAHUlNTBwgBVw==',
  `User-Agent` = 'Mozilla/5.0 (Macintosh; Intel Mac OS X 10_14_6) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/78.0.3904.87 Safari/537.36',
  `x-nba-stats-origin` = 'stats',
  `Sec-Fetch-Site` = 'same-origin',
  `Sec-Fetch-Mode` = 'cors',
  `Referer` = 'https://stats.nba.com/players/leaguedashplayerbiostats/',
  `Accept-Encoding` = 'gzip, deflate, br',
  `Accept-Language` = 'en-US,en;q=0.9'
)

url <- "https://stats.nba.com/stats/shotchartdetail?AheadBehind=&CFID=33&CFPARAMS=2019-20&ClutchTime=&Conference=&ContextFilter=&ContextMeasure=FGA&DateFrom=&DateTo=&Division=&EndPeriod=10&EndRange=28800&GROUP_ID=&GameEventID=&GameID=&GameSegment=&GroupID=&GroupMode=&GroupQuantity=5&LastNGames=0&LeagueID=00&Location=&Month=0&OnOff=&OpponentTeamID=0&Outcome=&PORound=0&Period=0&PlayerID=977&PlayerID1=&PlayerID2=&PlayerID3=&PlayerID4=&PlayerID5=&PlayerPosition=&PointDiff=&Position=&RangeType=0&RookieYear=&Season=2005-06&SeasonSegment=&SeasonType=Regular+Season&ShotClockRange=&StartPeriod=1&StartRange=0&StarterBench=&TeamID=0&VsConference=&VsDivision=&VsPlayerID1=&VsPlayerID2=&VsPlayerID3=&VsPlayerID4=&VsPlayerID5=&VsTeamID="

res <- httr::GET(url = url, httr::add_headers(.headers=headers))
json_resp <- jsonlite::fromJSON(content(res, "text"))

df <- data.frame(json_resp$resultSets$rowSet[1])
colnames(df) <- json_resp[["resultSets"]][["headers"]][[1]]

df$LOC_X <- as.numeric(as.character(df$LOC_X))
df$LOC_Y <- as.numeric(as.character(df$LOC_Y))

df %>% 
  ggplot(aes(x = LOC_X, y = LOC_Y)) + 
  geom_point() +
  coord_equal() + xlim(-200, 200) + ylim(0, 450)

kobe <- mean(as.numeric(df$SHOT_DISTANCE))
```


```{r}
ggplot(df, aes(x=LOC_X, y=LOC_Y) ) +
  geom_point() +
  theme_bw() + xlim(-250, 250) + ylim(-50, 450) + ggtitle("Kobe Bryant FGA: 2005 - 2006")
```



```{r}
headers = c(
  `Connection` = 'keep-alive',
  `Accept` = 'application/json, text/plain, */*',
  `x-nba-stats-token` = 'true',
  `X-NewRelic-ID` = 'VQECWF5UChAHUlNTBwgBVw==',
  `User-Agent` = 'Mozilla/5.0 (Macintosh; Intel Mac OS X 10_14_6) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/78.0.3904.87 Safari/537.36',
  `x-nba-stats-origin` = 'stats',
  `Sec-Fetch-Site` = 'same-origin',
  `Sec-Fetch-Mode` = 'cors',
  `Referer` = 'https://stats.nba.com/players/leaguedashplayerbiostats/',
  `Accept-Encoding` = 'gzip, deflate, br',
  `Accept-Language` = 'en-US,en;q=0.9'
)

url <- "https://stats.nba.com/stats/shotchartdetail?AheadBehind=&CFID=33&CFPARAMS=2019-20&ClutchTime=&Conference=&ContextFilter=&ContextMeasure=FGA&DateFrom=&DateTo=&Division=&EndPeriod=10&EndRange=28800&GROUP_ID=&GameEventID=&GameID=&GameSegment=&GroupID=&GroupMode=&GroupQuantity=5&LastNGames=0&LeagueID=00&Location=&Month=0&OnOff=&OpponentTeamID=0&Outcome=&PORound=0&Period=0&PlayerID=2544&PlayerID1=&PlayerID2=&PlayerID3=&PlayerID4=&PlayerID5=&PlayerPosition=&PointDiff=&Position=&RangeType=0&RookieYear=&Season=2009-10&SeasonSegment=&SeasonType=Regular+Season&ShotClockRange=&StartPeriod=1&StartRange=0&StarterBench=&TeamID=0&VsConference=&VsDivision=&VsPlayerID1=&VsPlayerID2=&VsPlayerID3=&VsPlayerID4=&VsPlayerID5=&VsTeamID="

res <- httr::GET(url = url, httr::add_headers(.headers=headers))
json_resp <- jsonlite::fromJSON(content(res, "text"))

df <- data.frame(json_resp$resultSets$rowSet[1])
colnames(df) <- json_resp[["resultSets"]][["headers"]][[1]]

df$LOC_X <- as.numeric(as.character(df$LOC_X))
df$LOC_Y <- as.numeric(as.character(df$LOC_Y))

df %>% 
  ggplot(aes(x = LOC_X, y = LOC_Y)) + 
  geom_point() +
  coord_equal() + xlim(-200, 200) + ylim(0, 450)

lebron<- mean(as.numeric(df$SHOT_DISTANCE))
```


```{r}
ggplot(df, aes(x=LOC_X, y=LOC_Y) ) +
  geom_point() +
  theme_bw() + xlim(-250, 250) + ylim(-50, 450) + ggtitle("LeBron James FGA: 2009 - 2010")
```



```{r}
headers = c(
  `Connection` = 'keep-alive',
  `Accept` = 'application/json, text/plain, */*',
  `x-nba-stats-token` = 'true',
  `X-NewRelic-ID` = 'VQECWF5UChAHUlNTBwgBVw==',
  `User-Agent` = 'Mozilla/5.0 (Macintosh; Intel Mac OS X 10_14_6) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/78.0.3904.87 Safari/537.36',
  `x-nba-stats-origin` = 'stats',
  `Sec-Fetch-Site` = 'same-origin',
  `Sec-Fetch-Mode` = 'cors',
  `Referer` = 'https://stats.nba.com/players/leaguedashplayerbiostats/',
  `Accept-Encoding` = 'gzip, deflate, br',
  `Accept-Language` = 'en-US,en;q=0.9'
)

url <- "https://stats.nba.com/stats/shotchartdetail?AheadBehind=&CFID=33&CFPARAMS=2019-20&ClutchTime=&Conference=&ContextFilter=&ContextMeasure=FGA&DateFrom=&DateTo=&Division=&EndPeriod=10&EndRange=28800&GROUP_ID=&GameEventID=&GameID=&GameSegment=&GroupID=&GroupMode=&GroupQuantity=5&LastNGames=0&LeagueID=00&Location=&Month=0&OnOff=&OpponentTeamID=0&Outcome=&PORound=0&Period=0&PlayerID=202695&PlayerID1=&PlayerID2=&PlayerID3=&PlayerID4=&PlayerID5=&PlayerPosition=&PointDiff=&Position=&RangeType=0&RookieYear=&Season=2019-20&SeasonSegment=&SeasonType=Regular+Season&ShotClockRange=&StartPeriod=1&StartRange=0&StarterBench=&TeamID=0&VsConference=&VsDivision=&VsPlayerID1=&VsPlayerID2=&VsPlayerID3=&VsPlayerID4=&VsPlayerID5=&VsTeamID="

res <- httr::GET(url = url, httr::add_headers(.headers=headers))
json_resp <- jsonlite::fromJSON(content(res, "text"))

df <- data.frame(json_resp$resultSets$rowSet[1])
colnames(df) <- json_resp[["resultSets"]][["headers"]][[1]]

df$LOC_X <- as.numeric(as.character(df$LOC_X))
df$LOC_Y <- as.numeric(as.character(df$LOC_Y))

df %>% 
  ggplot(aes(x = LOC_X, y = LOC_Y)) + 
  geom_point() +
  coord_equal() + xlim(-200, 200) + ylim(0, 450)

kawhi <- mean(as.numeric(df$SHOT_DISTANCE))
```


```{r}
ggplot(df, aes(x=LOC_X, y=LOC_Y) ) +
  geom_point() +
  theme_bw() + xlim(-250, 250) + ylim(-50, 450) + ggtitle("Kawhi Leonard FGA: 2019 - 2020")
```




```{r}
mj_advanced <- read_csv('jordanadvanced.csv')
str(mj_advanced) # seeing the data types of the variables
dim(mj_advanced) # checking the dimensions in order to subset columns in vars()
mj_advanced
mj_advanced_prime <- mj_advanced %>% filter(Age %in% c(27:29, 32:34)) 
mj_advanced_prime$PrimeYear <- 1:6
mj_advanced_prime
```



```{r}
dwade_advanced <- read_csv('wadeadvanced.csv')
dwade_advanced_prime <- dwade_advanced %>% filter(Age %in% c(24:25, 27:29, 31))
dwade_advanced_prime$PrimeYear <- 1:6
dwade_advanced_prime
```



```{r}
derozan_advanced <- read_csv('derozanadvanced.csv')
derozan_advanced_prime <- derozan_advanced %>% filter(Age %in% c(26:29, 32:33))
derozan_advanced_prime$PrimeYear <- 1:6
derozan_advanced_prime
```



```{r}
kobe_advanced <- read_csv('kobeadvanced.csv')
kobe_advanced_prime <- kobe_advanced %>% filter (Age %in% c(22:23, 27:28, 30:31))
kobe_advanced_prime$PrimeYear <- 1:6
kobe_advanced_prime
```



```{r}
lebron_advanced <- read_csv('lebronadvanced.csv')
lebron_advanced_prime <- lebron_advanced %>% filter (Age %in% c(24:29))
lebron_advanced_prime$PrimeYear <- 1:6
lebron_advanced_prime
```



```{r}
kawhi_advanced <- read_csv('kawhiadvanced.csv')
kawhi_advanced_prime <- kawhi_advanced %>% filter(Age %in% c(24:29))
kawhi_advanced_prime$PrimeYear <- 1:6
kawhi_advanced_prime
```



```{r}
ggplot() +
  geom_line(data = mj_advanced_prime, aes(x = PrimeYear, y = BPM), color = "red") + 
  geom_line(data = dwade_advanced_prime, aes(x = PrimeYear, y = BPM), color = "black") + 
  geom_line(data = derozan_advanced_prime, aes(x = PrimeYear, y = BPM, color = "gray")) + 
  geom_line(data = kobe_advanced_prime, aes(x = PrimeYear, y = BPM), color = "purple") + 
  geom_line(data = lebron_advanced_prime, aes(x = PrimeYear, y = BPM), color = "yellow") +
  geom_line(data = kawhi_advanced_prime, aes(x = PrimeYear, y = BPM), color = "blue") +
  geom_point(data = mj_advanced_prime, aes(x = PrimeYear, y = BPM), color = "red") + 
  geom_point(data = dwade_advanced_prime, aes(x = PrimeYear, y = BPM), color = "black") +
  geom_point(data = derozan_advanced_prime, aes(x = PrimeYear, y = BPM, color = "gray")) + 
  geom_point(data = kobe_advanced_prime, aes(x = PrimeYear, y = BPM), color = "purple") + 
  geom_point(data = lebron_advanced_prime, aes(x = PrimeYear, y = BPM), color = "yellow") +
  geom_point(data = kawhi_advanced_prime, aes(x = PrimeYear, y = BPM), color = "blue") +
  scale_color_manual(name='Player',
                     breaks=c('Jordan', 'Wade', 'DeRozan', 'Bryant', 'James', 'Leonard'),
                     values=c('Jordan' = "red", 'Wade' = "black", 'DeRozan' = "gray", 'Bryant'="purple", 'James'="yellow", 'Leonard'="blue")) +
  theme(legend.title=element_text(size=12), legend.text=element_text(size=10)) 
```


```{r}
abpm <- tribble(
~Player, ~`Avg_BPM`,
#Sad?/12 months/an account
"Michael Jordan", mean(mj_advanced_prime$BPM),
"Dwyane Wade", mean(dwade_advanced_prime$BPM),
"DeMar DeRozan", mean(derozan_advanced_prime$BPM),
"Kobe Bryant", mean(kobe_advanced_prime$BPM),
"LeBron James", mean(lebron_advanced_prime$BPM),
"Kawhi Leonard", mean(kawhi_advanced_prime$BPM),
)

abpm %>% ggplot() +
  geom_bar(aes(x = Player, y = `Avg_BPM`, color = Player, fill = Player), stat = 'identity')
```



```{r}
abpm$Avg_BPM - mean(mj_advanced_prime$BPM)
```



```{r}
ggplot() +
  geom_line(data = mj_advanced_prime, aes(x = PrimeYear, y = WS), color = "red") + 
  geom_line(data = dwade_advanced_prime, aes(x = PrimeYear, y = WS), color = "black") + 
  geom_line(data = derozan_advanced_prime, aes(x = PrimeYear, y = WS, color = "gray")) + 
  geom_line(data = kobe_advanced_prime, aes(x = PrimeYear, y = WS), color = "purple") + 
  geom_line(data = lebron_advanced_prime, aes(x = PrimeYear, y = WS), color = "yellow") +
  geom_line(data = kawhi_advanced_prime, aes(x = PrimeYear, y = WS), color = "blue") +
  geom_point(data = mj_advanced_prime, aes(x = PrimeYear, y = WS), color = "red") + 
  geom_point(data = dwade_advanced_prime, aes(x = PrimeYear, y = WS), color = "black") +
  geom_point(data = derozan_advanced_prime, aes(x = PrimeYear, y = WS, color = "gray")) + 
  geom_point(data = kobe_advanced_prime, aes(x = PrimeYear, y = WS), color = "purple") + 
  geom_point(data = lebron_advanced_prime, aes(x = PrimeYear, y = WS), color = "yellow") +
  geom_point(data = kawhi_advanced_prime, aes(x = PrimeYear, y = WS), color = "blue") +
  scale_color_manual(name='Player',
                     breaks=c('Jordan', 'Wade', 'DeRozan', 'Bryant', 'James', 'Leonard'),
                     values=c('Jordan' = "red", 'Wade' = "black", 'DeRozan' = "gray", 'Bryant'="purple", 'James'="yellow", 'Leonard'="blue")) +
  theme(legend.title=element_text(size=12), legend.text=element_text(size=10)) 
```


```{r}
aws <- tribble(
~Player, ~`Avg_WS`,
#Sad?/12 months/an account
"Michael Jordan", mean(mj_advanced_prime$WS),
"Dwyane Wade", mean(dwade_advanced_prime$WS),
"DeMar DeRozan", mean(derozan_advanced_prime$WS),
"Kobe Bryant", mean(kobe_advanced_prime$WS),
"LeBron James", mean(lebron_advanced_prime$WS),
"Kawhi Leonard", mean(kawhi_advanced_prime$WS),
)

aws %>% ggplot() +
  geom_bar(aes(x = Player, y = `Avg_WS`, color = Player, fill = Player), stat = 'identity')
```

```{r}
aws$Avg_WS - mean(mj_advanced_prime$WS)
```


```{r}
jordan - jordan
wade
derozan
kobe
lebron
kawhi
wade - jordan
derozan - jordan
kobe - jordan
lebron - jordan
kawhi - jordan
```

