---
title: "Data Journalism Spring 2023"
author: "Joshua Susanto"
date: "2023-05-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggridges)
library(ggplot2)
library(viridis)
library(hrbrthemes)
```

**Title:** How Can We Predict a Breakout Season?

**Idea:** Take a group of players who have had a “breakout season” (create a criteria that constitutes  what is considered a breakout season) and compare a variety of their stats, team composition, and other factors in order to determine a potential trend in having a breakout season.

**Method:**  We will select a group of about 25-30 players who have had a breakout season and gather their stats from both their breakout season as well as their previous seasons. We mean to look for trends in our data through focusing on a variety of these players stats including but not limited to:
Usage Ratings
Team composition
Basic Stats: PPG, RPG, APG, SPG, and shot percentages
Shot volume and shot selection
Box +/- : metric that estimates a player’s ability to impact the game, based on how they compare to an average players’ contribution to points per 100 possessions
Win Shares: tries to split up credit for the team’s successes to individual players
O/D Rating: offense/defense rating
Active Years
Position

We will need to focus on the differences in these stats between their prior and breakout season. Using different forms of analyses and visualizations, we will try to identify a certain trend or pattern between these players which may give us insight as to what goes into predicting a breakout season for an NBA player. 


**Sources:** 
https://www.basketball-reference.com/players/h/halibty01.html
https://www.basketball-reference.com/players/e/edwaran01.html
https://www.basketball-reference.com/players/g/grantje01.html
https://www.kaggle.com/datasets/drgilermo/nba-players-stats?resource=download&select=Seasons_Stats.csv


**What constitutes a breakout season**

- players have to had played a certain amount of time the prior season in order to account for injuries
  - at least 65 games to be considered for awards
  - for the sake of leniency, players must have played at least 55 games in the season
- use a statistic that best quantifies a player's overall value
  - according to multiple sources, the best stat for measuring a player's overall value is efficiency
- use the PER (player efficiency rating) in order to measure quantities of value
- see the largest jumps in PER, cut out data that has large jumps yet still below average or 60% percentile

Data used from kaggle: (https://www.kaggle.com/datasets/drgilermo/nba-players-stats?resource=download&select=Seasons_Stats.csv)

(note this data only has information up the 2017 NBA season)

```{r}
# import tidyverse
library(tidyverse)

# read in seasons data
seasons_data = read_csv('Seasons_Stats.csv')

head(seasons_data)
```


```{r}
# filter data, want player to have played over 50 games and want recorded PER, remove special characters from the player's name

seasons_data_clean = seasons_data %>% filter(!is.na(PER), PER <= 40, PER > 0, G > 50) %>% mutate(Player = str_replace_all(Player, "[^[:alnum:]]", " ")) %>% mutate(Player = str_trim(Player, side = 'right')) 
head(seasons_data_clean)
```




```{r}
# filter all players that have played more than one season 
elligible.players = seasons_data_clean %>% group_by(Player) %>% count() %>% filter(n > 1) 
players = elligible.players$Player
head(players, 20)
```



```{r}
# this function is used to calculate the largest per difference for every player
# this model fails to consider the possibility that multiple players can have the same name in the same season, can be improved
max_diff = c()
max_diff_season = c()
max_diff_per = c()
for (i in players) {
  temp = seasons_data_clean %>% filter(Player == i) 
  per = temp$PER
  per_diff = c(0)
  for (j in 1:(length(per)-1)) {
    diff = per[j+1] - per[j]
    per_diff = append(per_diff, diff)
    
  }
  max_diff = append(max_diff, max(per_diff))
  max_season = temp$Year[which(per_diff == max(per_diff))[1]] 
  max_diff_season = append(max_diff_season, max_season)
  temp2 = seasons_data_clean %>% filter(Player == i, Year == max_season) %>% dplyr::select(PER)
  max_diff_per = append(max_diff_per, temp2$PER[1])
}


```


Now that we have all of our players' biggest PER improvement, we can show our results in descending order to see which player/season have had the largest improvement.


```{r}
potential_breakout_seasons <- tibble(
  Player = players,
  `Year` = max_diff_season,
  `PER Differential` = max_diff,
  `PER` = max_diff_per
)
potential_breakout_seasons = potential_breakout_seasons %>% arrange(desc(`PER Differential`))
potential_breakout_seasons

```

```{r}
hist(potential_breakout_seasons$`PER Differential`, breaks = 50, main = 'Histogram of PER Differentials', xlab = 'PER Differential', col = 'lightgray')
```

Now we must further quantify a breakout season by imposing a minimum PER in order to be considered a breakout season. Our floor will be the average PER plus 3, which is always 18 by definition. 

```{r}
cutoff = potential_breakout_seasons$`PER Differential`[(1960-1960*0.8)] #finding 80th percentile of per diff

breakout_seasons = potential_breakout_seasons %>% filter(PER >= 19) %>% mutate(`Previous PER` = PER - `PER Differential`)
print(breakout_seasons)
```

And we are left with 395 potential breakout seasons in the entirety of NBA history (of available data). For more familiar names we can see the best breakout seasons since the 2000s

```{r}
hist(breakout_seasons$`PER Differential`, breaks = 50, main = 'Histogram of PER Differentials', xlab = 'PER Differential', col = 'lightgray')
```



```{r}
# must account for the potential season jump
modern = breakout_seasons %>% filter(Year >= 2000)
modern
```




```{r}
breakout_seasons2 = breakout_seasons %>% filter(`PER Differential` >= cutoff)

breakout_seasons2 # breakout seasons

hist(breakout_seasons2$`PER Differential`, breaks = 50, main = 'Histogram of PER Differentials', xlab = 'PER Differential', col = 'lightgray')
```


```{r}
breakout_players = breakout_seasons2$Player

seasons_data_clean %>% filter(Player %in% breakout_players)

# position, usage, and age

breakout3 = breakout_seasons2 %>% left_join(seasons_data_clean)


breakout3 %>% ggplot(aes(x=Pos, fill = Pos)) +  
  geom_bar( ) +
  scale_fill_hue(c = 40) +
  theme(legend.position="none") +
  ggtitle('Positions of NBA Breakout Seasons')
```

```{r}
ggplot(breakout3, aes(x = `Age`, y = `Pos`, fill = ..x..)) +
  geom_density_ridges_gradient(scale = 2, rel_min_height = 0.001) +
  scale_fill_viridis(name = "Temp. [F]", option = "C") +
  labs(title = 'Position vs Age for NBA Breakout Seasons') +
  theme_ipsum() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    )
```

```{r}
prev_team = c()
prev_mp = c()
prev_usg = c()
prev_ws = c()
prev_g = c()


# code to get all of our desired data for the previous seasons
for (i in 1:length(breakout3$Player)) {
   player = breakout3$Player[i]
   year = seasons_data_clean %>% filter(Player == player, PER == breakout3$`Previous PER`[i]) %>% select(Year)
   if (dim(year)[1] > 1) {
     year = year[1,1]
   } else {
    year = year[1] 
   }
   temp2 = seasons_data_clean %>% filter(Year == as.integer(year), Player == player)
   prev_team = append(prev_team, temp2$Tm[1])
   prev_mp = append(prev_mp, temp2$MP[1])
   prev_usg = append(prev_usg, temp2$`USG%`[1])
   prev_ws = append(prev_ws, temp2$WS[1])
   prev_g = append(prev_g, temp2$G[1])
}

breakout3$prev_team = prev_team
breakout3$prev_mp = prev_mp
breakout3$prev_usg = prev_usg
breakout3$prev_ws = prev_ws
breakout3$prev_g = prev_g

breakout3

sum(breakout3$Tm == breakout3$prev_team)

z = breakout3$Tm == breakout3$prev_team

team_change = ifelse(z == TRUE, "Yes", "No")

breakout3$team_change = team_change
```

```{r}

data <- data.frame(
  type = c(rep("Minutes Played (Prior)", 120), rep("Minutes Played (Breakout)", 120) ),
  value = c(breakout3$prev_mp, breakout3$MP)
)


a = data %>%
  ggplot(aes(x=value, fill=type)) +
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity', bins = 50) +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    theme_ipsum() +
    ggtitle('Histogram of Minutes Played') +
    labs(fill="") 


data2 <- data.frame(
  type = c(rep("Usage % (Prior)", 120), rep("Usage % (Breakout)", 120) ),
  value = c(breakout3$prev_usg, breakout3$`USG%`)
)


b = data2 %>%
  ggplot(aes(x=value, fill=type)) +
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity', bins = 50) +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    ggtitle('Histogram of Usage Percentage') +
    theme_ipsum() +
    labs(fill="") 

a
b
```

```{r}
changed = breakout3 %>% filter(team_change == 'Yes')
notchanged = breakout3 %>% filter(team_change == 'No')

ggplot(data=breakout3, aes(x=MP, group=team_change, fill=team_change)) +
    geom_density(adjust=1.5, alpha=.4) +
    ggtitle('Minutes Played Density Grouped by Team Change') +
    xlim(500,4000) 
    theme_ipsum()

ggplot(data=breakout3, aes(x=`USG%`, group=team_change, fill=team_change)) +
    geom_density(adjust=1.5, alpha=.4) +
    ggtitle('Usage Percentage Density Grouped by Team Change') +
    xlim(0,50) 
    theme_ipsum()
```

```{r}
data <- data.frame(
  type = c(rep("Minutes Played (Prior)", 87), rep("Minutes Played (Breakout)", 87) ),
  value = c(changed$prev_mp, changed$MP)
)


a = data %>%
  ggplot(aes(x=value, fill=type)) +
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity', bins = 50) +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    theme_ipsum() +
    ggtitle('Histogram of Minutes Played') +
    labs(fill="") 


data2 <- data.frame(
  type = c(rep("Usage % (Prior)", 120), rep("Usage % (Breakout)", 120) ),
  value = c(breakout3$prev_usg, breakout3$`USG%`)
)


b = data2 %>%
  ggplot(aes(x=value, fill=type)) +
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity', bins = 50) +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    ggtitle('Histogram of Usage Percentage') +
    theme_ipsum() +
    labs(fill="") 

a
b
```

```{r}
season23 = read_csv('season23advanced.csv')
```

```{r}
season23 %>% filter(G >= 50, Age < 25, Age > 21, PER > 13)
```

